### 2018-12-27 Sections/models_for_quantum_computation.tex
bib

\subsection{Adiabatic quantum computation \& quantum annealing} \label{sec:adiabatic_QC} \index{Adiabatic quantum computation} \index{Quantum annealing}

Adiabatic quantum computation (AQC) started as an approach to solving optimization problems, and has evolved into a universal alternative to the circuit model of quantum computing. 
AQC is based on an idea that is distinct from the circuit model \cite{RevModPhys.90.015002}. In the circuit model, the computation is encoded into a series of unitary quantum logic gates, whereas in AQC the computation starts from an initial whose ground state is easy to prepare, and evolves to a final state that encodes  the solution to the computation problem.

The adiabatic theorem forms the the backbone of AQC, which states that for a system initially prepared in an eigenstate of a time-dependent Hamiltonian $H(t)$, the evolution

\begin{align}
i \frac{\partial \ket{\psi(t)}}{\partial t} = H(t)\ket{\psi(t)}
\end{align}
\noindent will keep approximately to the instantaneous ground state, if $H(t)$ varies sufficiently slowly.

In the circuit model, the cost of the computation is equated with the number of gates. In AQC, Ahranov et al \cite{} defined the cost of adiabatic algorithm to be the dimensionless quantity

\begin{align}
\text{cost} = t_f \text{max}_s ||H(s)||
\end{align}
\noindent where $t_f$ is the run time, and $||\cdot||$ denotes the operator norm, which is the largest singular value of the operator.

The run time $t_f$ of an adiabatic algorithm scales at worst as $1/\Delta^3$, 
where $\Delta^3$ is the minimum eigenvalue gap between the ground state and the first excited state of the Hamiltonian\cite{}. This is the reason why AQC has generated much interest: it has a rich connection to well-studied condense matter physics.

The analysis for AQC involve bounding the eigenvalue gap of a complicated many-body Hamiltonian, which is a notoriously difficult problem. Nevertheless, a number of examples are known where a gap between classical and quantum. 
These include:

\emph{The adiabatic quantum Grover algorithm} - as its name suggest, this is the adiabatic version of Grover's search algorithm. The initial Hamiltonian is $H_0 = \openone - \ket{\phi}\bra{\phi}$, where $\ket{\phi}$ is the univorm superposition state $ \ket{+}^{\otimes n}. $
The final Hamiltonian is given by $H_1 = \openone - \ket{m}\bra{m}$, where $ \ket{m}$ is the marked item we are searching for \cite{PhysRevA.65.042308}.

\emph{Adiabatic Deutsch-Joza algorithm} - Given a function $f:\{0,1 \}^n \rightarrow \{0,1\}$ which is either constant or balanced, the Deutsch-Joza algorithm can determine which type of function it is with a quantum speedup. An adiabatic implementation \cite{PhysRevLett.95.250503} is derived to match the speedup obtained as in the circuit model.

\emph{The glued-trees problem} - here we consider two binary trees each of depth $n$. Each tree has $2^{n+1}-1$ vertices, the two trees are randomly glued as in Fig~\ref{fig:glued_tree}. More specifically, a leaf is chosen from the left, is connected to a leaf on the right at random, which is then connected to another leaf on the left, and so on. Every leaf on the left is connected to two on the right and vice versa. This creats a random cycle that alternates between the two trees. The problem is to, starting from the left root, to find a path to the right root in the smallest possible number of steps. Classical algorithms require at least a subexponential in $n$ number of quiries, but there exists a polynomial-time quantum algorithm based on quantum walks \cite{childs2003exponential}.

\emph{Adiabatic PAGERANK algorithm} - the PAGERANK vector is a crucial tool in data mining and information retrival, used by the Google search engine. The current algorithm outperforms all known current classical algorithms by a polynomial or even exponential speedup \cite{PhysRevLett.108.230506}, but better future classical algorithms have not been ruled out.

\begin{figure}[!htbp]
%	\includegraphics[trim = 0 0 0 1.3cm, clip=true, width=0.5\textwidth]{glued_tree}
	\captionspacefig \caption{\comment{Missing PDF}A glued tree with $n=4$} \label{fig:glued_tree}
\end{figure}

The computational power of the circuit model and the adiabatic model of quantum computing are equivalent up to a polynomial overhead. In the circuit model, a set of gates is said to be universal if any unitary operation may be approximated to arbitrary accuracy by a quantum circuit involving only those gates. The analog of such a set of gates in AQC is a Hamiltonian.

A time-dependent Hamiltonian H(t) is universal if, given an arbitrary quantum circuit $U$ operating on an arbitrary multipartite initial state $\ket{\psi}$, the ground state of $H(t_f)$ is equal to $U\ket{\psi}$.

It has been proven that AQC can efficiently simulate the circuit model \cite{RevModPhys.90.015002}, which means that it is universal for quantum computing.

>>>>

\subsection{Adiabatic quantum computation \& quantum annealing} \label{sec:adiabatic_QC} \index{Adiabatic!Quantum computation} \index{Quantum annealing}

Adiabatic quantum computation (AQC) began as an approach for solving optimisation problems, but has evolved into a universal alternative to the circuit model for quantum computing. AQC is based on an idea that is distinct from the circuit model \cite{bib:RevModPhys.90.015002}. In the circuit model, the computation is encoded into a series of unitary quantum logic gates, whereas in AQC the computation starts from an initial Hamiltonian whose ground state is easy to prepare, and evolves to a final state that encodes the solution to the computational problem.

The adiabatic theorem\index{Adiabatic!Theorem} forms the backbone of AQC, which states that for a system initially prepared in an eigenstate of a time-dependent Hamiltonian $\hat{H}(t)$, the evolution via the Schr{\"o}dinger equation\index{Schr{\"o}dinger equation},
\begin{align}
i \frac{\partial \ket{\psi(t)}}{\partial t} = \hat{H}(t)\ket{\psi(t)},
\end{align}
will keep approximately to the instantaneous ground state, if $\hat{H}(t)$ varies sufficiently slowly.

In the circuit model, the cost of the computation is equated with the number of gates. In AQC, \cite{???} defined the cost of adiabatic algorithms to be the dimensionless quantity,
\begin{align}
\mathrm{cost} = t_f \cdot \mathrm{max}_s ||\hat{H}(s)||,
\end{align}
where $t_f$ is the algorithmic runtime, and $||\cdot||$ denotes the operator norm\index{Operator norm}, its largest singular value\index{Singular value}.

The runtime $t_f$ of an adiabatic algorithm scales at worst as \mbox{$O(1/\Delta^3)$}, where $\Delta^3$ is the minimum eigenvalue gap between the ground state and the first excited state of the Hamiltonian \cite{???}. This is the reason why AQC has generated much interest -- it has a rich connection to the well-studied field of condensed matter physics\index{Condensed matter physics}.

The analysis for AQC involves bounding the eigenvalue gap of a complicated many-body Hamiltonian, which is a notoriously difficult problem. Nevertheless, a number of examples are known exhibiting a gap between the classical and quantum cases.

Examples of adiabatic quantum algorithms include:
\begin{itemize}
\item Adiabatic Grover algorithm\index{Adiabatic!Grover algorithm}: as its name suggests, this is the adiabatic version of Grover's search algorithm. The initial Hamiltonian is,
\begin{align}
	\hat{H}_0 = \hat\openone - \ket{\phi}\bra{\phi},
\end{align}
where $\ket{\phi}$ is the uniform superposition state $\ket{+}^{\otimes n}$. The final Hamiltonian is given by,
\begin{align}
\hat{H}_1 = \hat\openone - \ket{m}\bra{m},
\end{align}
where $\ket{m}$ is the marked item we are searching for \cite{bib:PhysRevA.65.042308}.
\item Adiabatic Deutsch-Jozsa algorithm\index{Adiabatic!Deutsch-Jozsa algorithm}\index{Deutsch-Jozsa algorithm}: given a binary function,
\begin{align}
f:\{0,1 \}^n \rightarrow \{0,1\},
\end{align}
which is either constant\index{Constant functions} or balanced\index{Balanced functions}, the Deutsch-Jozsa algorithm can determine which type of function it is, exhibiting quantum speedup. An adiabatic implementation \cite{bib:PhysRevLett.95.250503} is derived to match the speedup obtained in the original circuit model implementation.
\item Adiabatic glued-trees problem\index{Adiabatic!Glued-trees problem}: here we consider two binary trees\index{Binary trees} of depth $n$. Each tree has \mbox{$2^{n+1}-1$} vertices, and the two trees are randomly glued together, as shown in Fig.~\ref{fig:glued_tree}. More specifically, a leaf is chosen from the left, is connected to a leaf on the right also chosen at random, which is then connected to another leaf on the left, and so on. Every leaf on the left is connected to two on the right and vice versa. This creates a random cycle that alternates between the two trees. The problem is to, starting from the left root, find a path to the right root in the smallest possible number of steps. Classical algorithms require at least a sub-exponential number of queries, but there exists a polynomial-time quantum algorithm based on quantum walks \cite{bib:childs2003exponential}.
\item Adiabatic \textsc{PageRank} algorithm\index{Adiabatic!PageRank algorithm}: the \textsc{PageRank}\index{PageRank algorithm} vector is a crucial tool in data mining and information retrieval, employed heavily by the Google search engine. The goal of the algorithm is to rank the importance, impact or influence of some entity in a network (webpages on the internet in the case of a Google search). The algorithm achieves this by representing the network as a flow network\index{Flow network}, whereby flow from one vertex to another represents a vote by the source vertex for the importance of the destination vertex. By finding a steady-state flow to each edge in the network, the net flow into vertices represents their cumulative importance, as determined collectively by participants in the network. However, the magnitude of votes cast by vertices is weighted by their own ranking. Therefore, the algorithm aims to filter out bogus gaming of the system via the creation of dummy nodes that cast votes en masse to rig the election. The \textsc{PageRank} approach to ranking webpages has been by far the most successful and robust such algorithm, and was perhaps the key development behind the global success of Google. The best-known classical algorithms for solving the \textsc{PageRank} problem are via: representing the flow network as a matrix equation, from which the solution is obtained via solving an eigenvalue equation; representing the flow network as a random walk, whereby walkers randomly follow paths influenced by the flows, which in the long-time limit yields walks approximating the solution. The current best-known quantum \textsc{PageRank} algorithm outperforms all currently known classical algorithms with polynomial or even exponential speedup \cite{bib:PhysRevLett.108.230506}, but better future classical algorithms have not been ruled out and are under active investigation.
\end{itemize}

\begin{figure}[!htbp]
%	\includegraphics[trim = 0 0 0 1.3cm, clip=true, width=0.5\textwidth]{glued_tree}
	\captionspacefig \caption{\comment{Missing PDF}A glued tree with \mbox{$n=4$}.} \label{fig:glued_tree}
\end{figure}

The computational power of the circuit model and the adiabatic model for quantum computing are equivalent up to a polynomial overhead. In the circuit model, a set of gates is said to be universal if any unitary operation may be approximated to arbitrary accuracy by a quantum circuit involving only those gates. The analog of such a set of gates in AQC is a Hamiltonian.

A time-dependent Hamiltonian $\hat{H}(t)$ is universal if, given an arbitrary quantum circuit $\hat{U}$ operating on an arbitrary multipartite initial state $\ket{\psi}$, the ground state of $\hat{H}(t_f)$ is equal to $\hat{U}\ket{\psi}$.

It has been proven that AQC can efficiently simulate the entire circuit model \cite{bib:RevModPhys.90.015002}, which implies that it is itself universal for efficient quantum computing.
